Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-08-03 paper/paper.Rmd
revised all figure captions; added a few TODO comments

```{r setup-2, echo=FALSE, purl=T}
# for knitr spin later
knitr::opts_chunk$set(fig.path = "./paper/figs/",
  cache = FALSE,
  cache.lazy = TRUE)
```

@@

```{r disc, fig.cap="Visualization of 'Leave-one-disc-out' method.", out.width="40%"}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
resampling_disc = rsmp("spcv_disc", folds = 5, radius = 300L, buffer = 400L)
resampling_disc$instantiate(task)

@@

```{r buffer, fig.cap="Visualization of the 'Spatial buffering' method from package `blockCV`.", out.width="40%", fig.pos="ht"}
resampling_buffer = rsmp("spcv_buffer", theRange = 1000)
resampling_buffer$instantiate(task)

@@

```{r block-random, fig.cap="Spatial blocking with option `selection = 'random'`.", out.width="40%", fig.pos="ht"}
resampling_block_random = rsmp("spcv_block", range = 1000, folds = 5)

@@

```{r block-systematic, fig.cap="Spatial blocking with option `selection = 'systematic'`.", out.width="40%", fig.pos="ht"}
resampling_block_systematic = rsmp("spcv_block",
  range = 1000, folds = 5,
  selection = "systematic"
)

@@

```{r block-checkerboard, fig.cap="Spatial blocking with option `selection = 'checkerboard'`.", out.width="40%", fig.pos="ht!"}
resampling_block_checkerboard = rsmp("spcv_block",
  range = 1000, folds = 5,
  selection = "checkerboard")

@@

```{r block-rows-and-cols, fig.cap="Spatial blocking with arguments `rows` and `cols`.", out.width="40%", fig.pos="ht"}
resampling_block_rows_cols = rsmp("spcv_block",
  rows = 6, cols = 3, folds = 5
)

@@

```{r tile, fig.cap="Spatial tiling", out.width="40%", fig.pos="ht"}
resampling_tiles = rsmp("spcv_tiles", nsplit = c(3L, 4L))

@@

```{r coords, fig.cap="Method 'Spatial CV' showing $k$-means clustering based on coordinates.", out.width="40%", fig.pos="ht"}
resampling_coords = rsmp("spcv_coords", folds = 5)$instantiate(task)

@@

```{r lto, echo=FALSE, out.width="70%", fig.cap="Three-dimensional visualization of method `sptcv\\_cstf` and sub-method Leave-Time-Out (LTO). To support showcasing the temporal grouping, five folds and five temporal levels in variable `Date` were used in this example. Due to the tilt angle of the static 3D plot the last two levels are hard to distinguish visually but they are separated in the same way as the the first two levels on the top of the plot.", fig.pos="ht"}
knitr::include_graphics("lto.png")
```

@@

```{r llo, echo=FALSE, out.width="70%", fig.cap="Three-dimensional visualization of method 'sptcv\\_cstf' and sub-method 'Leave-Location-Out' (LLO). To support showcasing the spatial grouping, the viewing angle has been adjusted to a view from top. Points are stacked by temporal levels and only points from the same location are chosen to form a partition.", fig.pos="ht"}
knitr::include_graphics("llo.png")
```

@@

```{r llto, echo=FALSE, out.width="70%", fig.cap="Three-dimensional visualization of method 'sptcv\\_cstf' and sub-method 'Leave-Location-And-Time-Out' (LLTO). The grey points show the points omitted by this method, i.e. the points which are neither used in the training nor test set in this fold.", fig.pos="ht"}
knitr::include_graphics("llto.png")
```

@@

```{r custom-cv, fig.cap="Custom resampling using an external factor vector defining the partitions. Each coloring represents one partition.", out.width="40%", fig.pos="ht"}
breaks = quantile(task$data()$dem, seq(0, 1, length = 6))
zclass = cut(task$data()$dem, breaks, include.lowest = TRUE)

@@

autoplot(resampling_custom, size = 0.8, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

```{r env-1, fig.cap="Environmental blocking using a single predictor.", out.width="40%", fig.pos="ht"}
resampling_env = rsmp("spcv_env",
  features = "distdeforest", folds = 5)

@@

```{r env-2, fig.cap="Environmental blocking using multiple predictors.", out.width="40%", fig.pos="ht"}
resampling_env_multi = rsmp("spcv_env",
  features = c("distdeforest", "slope"), folds = 5)

autoplot(resampling_env_multi,
  size = 0.8, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```

@@

```{r vis-spcv, fig.cap="Visualization of spatial partitioning using coordinate-based clustering. Due to space limitations only the first two folds are shown.", fig.pos="ht", fig.height=6, fig.width=8}
resampling_sp = rsmp("repeated_spcv_coords", folds = 4, repeats = 2)
resampling_sp$instantiate(task)

@@

```{r vis-nspcv, fig.cap="Visualization of random partitioning", fig.pos="ht", fig.height=6, fig.width=8}
resampling_nsp = rsmp("repeated_cv", folds = 4, repeats = 2)
resampling_nsp$instantiate(task)

>>>>

```{r setup-2, echo=FALSE, purl=T}
# for knitr spin later
knitr::opts_chunk$set(fig.path = "./paper/figs/",
  cache = TRUE,
  cache.lazy = TRUE)
```

@@

```{r disc, fig.cap='Visualization of one training set / test set combination generated with the leave-one-disc-out method from package sperrorest (method "spcv\\_disc" in mlr3spatiotempcv). The disc has a radius of 300 m and is surrounded by a 400-m buffer.', out.width="40%"}
library("mlr3")
library("mlr3spatiotempcv")
task = tsk("ecuador")
resampling_disc = rsmp("spcv_disc", folds = 5, radius = 300L, buffer = 400L)
resampling_disc$instantiate(task)

@@

```{r buffer, fig.cap='Visualization of the spatial buffering method from package blockCV (method "spcv\\_buffer" in  mlr3spatiotempcv). The buffer distance is 1000 m.', out.width="40%", fig.pos="ht"}
resampling_buffer = rsmp("spcv_buffer", theRange = 1000)
resampling_buffer$instantiate(task)

@@

```{r block-random, fig.cap='Random resampling of square spatial blocks using the implementation in package blockCV (method "spcv\\_block" with option selection = "random" in mlr3spatiotempcv). The size of the squares is 1000 m, and four out of the 19 blocks were assigned to this test sample.', out.width="40%", fig.pos="ht"}
resampling_block_random = rsmp("spcv_block", range = 1000, folds = 5)

@@

```{r block-systematic, fig.cap='Sytematic resampling of square spatial blocks using the implementation in package blockCV (method "spcv\\_block" with option selection = "systematic" in mlr3spatiotempcv). The size of the squares is 1000 m, and four out of the 19 blocks were assigned to this test sample.', out.width="40%", fig.pos="ht"}
resampling_block_systematic = rsmp("spcv_block",
  range = 1000, folds = 5,
  selection = "systematic"
)

@@

```{r block-checkerboard, fig.cap='Checkerboard resampling of square spatial blocks using the implementation in package blockCV (method "spcv\\_block" with option selection = "checkerboard" in mlr3spatiotempcv). The size of the squares is 1000 m, and eleven out of the 19 blocks were assigned to this test sample.', out.width="40%", fig.pos="ht!"}
resampling_block_checkerboard = rsmp("spcv_block",
  range = 1000, folds = 5,
  selection = "checkerboard")

@@

```{r block-rows-and-cols, fig.cap='Random resampling of a specifid number of rectangular blocks as implemented in package blockCV (method "spcv\\_block" in mlr3spatiotempcv with arguments rows = 6 and cols = 3).', out.width="40%", fig.pos="ht"}
resampling_block_rows_cols = rsmp("spcv_block",
  rows = 6, cols = 3, folds = 5
)

@@

```{r tile, fig.cap='Leave-one-block-out resampling from package sperrorest (method "spcv\\_tiles" in package mlr3spatiotempcv with argument nsplit = c(3,4) indicating the number of rows and columns).', out.width="40%", fig.pos="ht"}
resampling_tiles = rsmp("spcv_tiles", nsplit = c(3L, 4L))

@@

```{r coords, fig.cap='Leave-one-block-out CV based on $k$-means clustering of the coordinates as implemented in package sperrorest (method "spcv\\_coords" in mlr3spatiotempcv).', out.width="40%", fig.pos="ht"}
resampling_coords = rsmp("spcv_coords", folds = 5)$instantiate(task)

@@

```{r lto, echo=FALSE, out.width="70%", fig.cap='Perspective plot of leave-time-out CV from package CAST (method "sptcv\\_cstf" and argument time\\_var = "Date"). Only five folds and five time points were used in this example. Note that the blue dots correspond to five discrete time levels, which appear as a point cloud due to the viewing angle.', fig.pos="ht"}
knitr::include_graphics("lto.png")
```

@@

```{r llo, echo=FALSE, out.width="70%", fig.cap='Perspective plot of leave-location-out CV from package CAST (method "sptcv\\_cstf" and argument space\\_var = "SOURCEID"). The data is viewed from the future towards the past.', fig.pos="ht"}
knitr::include_graphics("llo.png")
```

@@

```{r llto, echo=FALSE, out.width="70%", fig.cap='Perspective plot of leave-location-and-time-out CV from package CAST (method "sptcv\\_cstf" and arguments time\\_var = "Date" and space\\_var = "SOURCEID"). The grey points are excluded from both the training and the test set in this fold.', fig.pos="ht"}
knitr::include_graphics("llto.png")
```

@@

<!-- TODO: AB: This diagram is inconsistent with the other diagrams - should display training and test samples. -->
```{r custom-cv, fig.cap='Leave-one-level-out (custom) resampling from package mlr3 (method "custom\\_cv"). A factor variable is used to define the partitions. Each color represents one partition.', out.width="40%", fig.pos="ht"}
breaks = quantile(task$data()$dem, seq(0, 1, length = 6))
zclass = cut(task$data()$dem, breaks, include.lowest = TRUE)

@@

autoplot(resampling_custom, size = 0.8, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))

```

@@

```{r env-1, fig.cap='Environmental leave-one-block-out CV from package blockCV using a single predictor to define blocks in feature space (method "spcv\\_env" in mlr3spatiotempcv with argument features = "distdeforest").', out.width="40%", fig.pos="ht"}
resampling_env = rsmp("spcv_env",
  features = "distdeforest", folds = 5)

@@

<!-- TODO: AB: should be combined with the previous figure into a two-panel figure. Differences are much easier to recognize if plotted next to each other. -->
```{r env-2, fig.cap='Environmental leave-one-block-out CV from package blockCV using two predictors to define blocks in feature space (method "spcv\\_env" in mlr3spatiotempcv with argument features = c("distdeforest", "slope")).', out.width="40%", fig.pos="ht"}
resampling_env_multi = rsmp("spcv_env",
  features = c("distdeforest", "slope"), folds = 5)

autoplot(resampling_env_multi,
  size = 0.8, fold_id = 1, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
```
<!-- AB: TODO: In general, the code would look much simpler if we could get rid of the scale_x/y_continuous() calls. Improve default settings of autoplot()? -->

@@

```{r vis-spcv, fig.cap="Spatial leave-one-block-out partitioning using coordinate-based clustering to create roughly equally sized polygonal blocks. Due to space limitations only the first two folds of the first repetition are shown.", fig.pos="ht", fig.height=6, fig.width=8}
resampling_sp = rsmp("repeated_spcv_coords", folds = 4, repeats = 2)
resampling_sp$instantiate(task)

@@

```{r vis-nspcv, fig.cap="Random (non-spatial) four-fold CV partitioning. Only the first two folds of the first repetition are shown.", fig.pos="ht", fig.height=6, fig.width=8}
resampling_nsp = rsmp("repeated_cv", folds = 4, repeats = 2)
resampling_nsp$instantiate(task)
